---
title: "The Solution Lab Test"
author: "Guojing Wu"
date: "5/22/2020"
output:
    pdf_document:
    highlight: default
    number_sections: true
    citation_package:
    keep_tex: false
    fig_caption: true
    latex_engine: pdflatex
fontsize: 10pt
geometry: margin=1in
bibliography:
biblio-style:
header-includes:
- \usepackage{indentfirst}
- \usepackage{graphicx}
- \usepackage{geometry}
- \usepackage{subfigure}
- \usepackage{amsmath}
- \usepackage{listings}
- \usepackage{tikz}
- \usetikzlibrary{matrix}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, comment = "")
library(dplyr)
library(Seurat)
library(patchwork)

```

## Load and Preprocess

```{r}
# Load the PBMC dataset
pbmc.data <- Read10X(data.dir = "filtered_gene_bc_matrices/hg19/")

# Initialize the Seurat object with the raw (non-normalized data).
pbmc <- CreateSeuratObject(counts = pbmc.data, project = "pbmc3k", min.cells = 3, min.features = 200)
pbmc

```

visualize QC metrics, and use these to filter cells:

* We filter cells that have unique feature counts over 2,500 or less than 200

* We filter cells that have >5% mitochondrial counts

```{r, dpi = 300, fig.width = 10}
# The [[ operator can add columns to object metadata. This is a great place to stash QC stats
pbmc[["percent.mt"]] <- PercentageFeatureSet(pbmc, pattern = "^MT-")

# Visualize QC metrics as a violin plot
VlnPlot(pbmc, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

```

Feature-feature relationships:

```{r, dpi = 300, fig.width = 10}
# FeatureScatter is typically used to visualize feature-feature relationships, but can be used
# for anything calculated by the object, i.e. columns in object metadata, PC scores etc.
plot1 <- FeatureScatter(pbmc, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(pbmc, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1 + plot2

```

Feature selection (2000 features):

```{r, dpi = 300, fig.width = 10}
# filtering and normalizing
pbmc <- subset(pbmc, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & percent.mt < 5)
pbmc <- NormalizeData(pbmc, normalization.method = "LogNormalize", scale.factor = 10000)

# feature selection
pbmc <- FindVariableFeatures(pbmc, selection.method = "vst", nfeatures = 2000)
# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(pbmc), 10)
# plot variable features with and without labels
plot1 <- VariableFeaturePlot(pbmc)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 + plot2

```

```{r}
# scaling
all.genes <- rownames(pbmc)
pbmc <- ScaleData(pbmc, features = all.genes)

```

## Linear Dimensional Reduction: PCA

```{r}
pbmc <- RunPCA(pbmc, features = VariableFeatures(object = pbmc))
# Examine and visualize PCA results a few different ways
print(pbmc[["pca"]], dims = 1:5, nfeatures = 5)

```

```{r, dpi = 300, fig.width = 10, fig.height = 6}
VizDimLoadings(pbmc, dims = 1:2, reduction = "pca")

```

```{r, dpi = 300}
DimPlot(pbmc, reduction = "pca")

```

Heatmap based on PCs:

```{r, dpi = 300, fig.width = 10, fig.height = 15}
# heatmap
DimHeatmap(pbmc, dims = 1:15, cells = 500, balanced = TRUE)

```

Determine the dimensionality:

* JackStraw plot:

```{r, dpi = 300}
# JackStraw
pbmc <- JackStraw(pbmc, num.replicate = 100)
pbmc <- ScoreJackStraw(pbmc, dims = 1:20)
JackStrawPlot(pbmc, dims = 1:15)

```

* Elbow plot:

```{r, dpi = 300}
ElbowPlot(pbmc)
```

Based on the plot, we roughly choose the first 10 PCs

## Clustering

Visualizing use UMAP:

```{r, dpi = 300}
# clustering
pbmc <- FindNeighbors(pbmc, dims = 1:10)
pbmc <- FindClusters(pbmc, resolution = 0.5)

# visualize result use UMAP
pbmc <- RunUMAP(pbmc, dims = 1:10)
DimPlot(pbmc, reduction = "umap")
```

```{r}
# save files
saveRDS(pbmc, file = "pbmc_tutorial.rds")

```

## Finding differentially expressed features

Markers for every cluster compared to all remaining cells:

```{r}
# find markers for every cluster compared to all remaining cells, report only the positive ones
pbmc.markers = FindAllMarkers(pbmc, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
pbmc.markers %>% 
  group_by(cluster) %>% 
  top_n(n = 2, wt = avg_logFC) %>% 
  knitr::kable()

```

Visualize using VlnPlot:

```{r, dpi = 300, fig.width = 10, fig.height = 6}
# Visualize using VlnPlot
VlnPlot(pbmc, features = c("MS4A1", "CD79A"))

```

```{r, dpi = 300, fig.width = 10, fig.height = 6}
# plot raw counts as well
VlnPlot(pbmc, features = c("NKG7", "PF4"), slot = "counts", log = TRUE)

```

Visualize using FeaturePlot:

```{r, dpi = 300, fig.width = 10, fig.height = 8}
FeaturePlot(pbmc, features = c("MS4A1", "GNLY", "CD3E", "CD14", "FCER1A", "FCGR3A", "LYZ", "PPBP", 
    "CD8A"))
```

Plotting the top 20 markers (or all markers if less than 20) for each cluster use DoHeatmap:

```{r, dpi = 300, fig.width = 10}
# plotting the top 20 markers (or all markers if less than 20) for each cluster
top10 = pbmc.markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_logFC)
DoHeatmap(pbmc, features = top10$gene) + NoLegend()

```

## Assigning cell type identity to clusters

```{r, dpi = 300}
# assigning cell type identity to clusters
new.cluster.ids = c("Naive CD4 T", "Memory CD4 T", "CD14+ Mono", "B", "CD8 T", "FCGR3A+ Mono", 
    "NK", "DC", "Platelet")
names(new.cluster.ids) = levels(pbmc)
pbmc = RenameIdents(pbmc, new.cluster.ids)
DimPlot(pbmc, reduction = "umap", label = TRUE, pt.size = 0.5) + NoLegend()

```

```{r}
saveRDS(pbmc, file = "pbmc3k_final.rds")
```

## Appendix A
```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
```
